\documentclass[11pt]{gsasthesis} % 10,11 and 12pt fonts allowed

%%%%%%%%%%%%%%%% PACKAGES YOU PROBABLY WANT %%%%%%%%%%%%%%%%
% Include packages you want. The gsasthesis style file already includes
% packages "setspace" and "tocbibind".

\usepackage{etex} % extend the number of registers

% GSAS: "all margins should be at least 1 inch."
\usepackage[margin={1.2in}]{geometry}
\usepackage{bbold}

% If you want asymmetric margins for two-sided documents, use the "twoside"
% option, as in
% \usepackage[top=1in,bottom=1.5in,left=1in,right=1.5in,twoside]{geometry} The
% left and right options become inner and outer margins The default horizontal
% latex margin ratio is 2:3. The default vertical top:bottom margin ratio is 2:3
% also. You can also set it directly by passing the hmarginratio option to the
% geometry package, as in
% \usepackage[top=1in,left=1in,vmarginratio=2:3,hmarginratio=2:5,twoside]{geometry}

% Appendix package. Not necessary, but it does make managing appendices easier
\usepackage[titletoc]{appendix}

%%%%%%%%%%%%%%%% PACKAGES MAY WANT %%%%%%%%%%%%%%%%

% sideways tables and figures
\usepackage{rotating}

% tables that spill over multiple pages
\usepackage{longtable}

% references
\usepackage{natbib}

% fonts that are nicer than defaults
\usepackage[sc]{mathpazo}
\usepackage{courier}

% Use 8-bit encoding that has 256 glyphs, pretty please
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% babel is required for blindtext, which generates random text
\usepackage[english]{babel}
\usepackage{blindtext}

% math support
\usepackage{amsmath}

% Slightly tweak font spacing for aesthetics
\usepackage{microtype}

% You need the footmisc package with the stable option if you want to have
% footnotes inside section titles, for example to say that a particular chapter
% has been co-authored with someone. The multiple option ensures that there is a
% comma between two consecutive footnotes
\usepackage[stable,multiple]{footmisc}

\usepackage{chngcntr}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\graphicspath{{./images/}{../images}}
\usepackage{algorithmicx}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{xcolor,url}
\usepackage[vlined,linesnumbered,ruled,resetcount]{algorithm2e}
\newcommand{\nextnr}{\stepcounter{AlgoLine}\ShowLn}
\usepackage{xr}

\counterwithout{footnote}{chapter}

% Nicer captions
\RequirePackage[font=small,format=plain,labelfont=bf,textfont=it]{caption}
\addtolength{\abovecaptionskip}{1ex}
\addtolength{\belowcaptionskip}{1ex}
\setlength{\parskip}{0.5cm plus2mm minus2mm}

%%%%%%%%%%%%%%%% COMPULSORY FIELDS %%%%%%%%%%%%%%%%

\title{Change Point Modeling and Regime Learning in Markov Switching Systems \\
\large Modeling the Effects of Students' Interactions with Immersive Simulations} % needs to match title on DAC
\author{Nicholas S. Hoernle} % full name as it appears on your GSAS record, needs
                          % to match name on DAC
\degreename{Master of Engineering}
\degreefield{Computational Science and Engineering}
                                % handbook
\department{Institute for Applied Computational Sciences} % official name of department
\degreemonth{May} % Month of Defense (i.e. month when DAC was signed)
\degreeyear{2018} % Year the DAC was signed
\principaladvisor{Dr. Pavlos Protopapas}

% Optionally, you can add a second advisor, but you can't have three
\secondadvisor{Prof. Barbara Grosz}



\begin{document}

%%%%%%%%%%%%%%%% FRONTMATTER %%%%%%%%%%%%%%%%

\pagenumbering{roman} % GSAS wants roman page numbers for frontmatter

% the following four pages are required in that order. The first two pages are
% not allowed to have page numbers, this is taken care of in the class file.
\thesistitlepage
% \copyrightpage
\begin{abstract}
Simulations that combine real world components with interactive digital media provide a rich setting for students with the potential to assist knowledge building and understanding of complex physical processes. This thesis addresses the problem of modeling the effects of multiple students' simultaneous interactions on the complex and exploratory environments such simulations provide. We work towards assisting educators with the difficult task of interpreting student exploration and we direct this research by considering the specific case of Connected Worlds, an innovative mixed reality simulation.

Connected Worlds is introduced to present a case study that motivates the need for providing assistive tools for teachers when working with complex exploratory learning simulations. The multi-person and immersive setting means that it is hard for the teacher to track the session and to identify salient learning opportunities. We aim to remove this burden from the teacher by extracting information from the log files of the simulation. Our goal is to find information in the logs that may be useful to a teacher when reviewing a specific session with a class of students.

We introduce switch based models and propose the switching state space model as a good candidate for decomposing the time series that is generated from a session with Connected Worlds. The model breaks the time series into periods that individually are easily interpreted. This work presents an inference algorithm that learns the transition points between successive periods in the time series as well as the internal dynamics that govern each period. This algorithm differs from other switch based inference schemes in that its primary goal is to decompose the time series in a way that is human interpretable.

The model was applied to data that was obtained from Connected Worlds and it was used to segment different sessions into multiple periods. Inferred parameters associated with the individual periods were used to generate text-based descriptions of the system dynamics. Two experimental visualizations of the model were designed. These visualizations included the generated text-based descriptions and a movie representation of the system dynamics. The experiments aimed to test the interpretability of the inferred periods and the text descriptions of the dynamics that were present alongside the associated video clip. Our results indicate that the switch based model finds relevant boundaries between salient periods and it generates representative text that describes the period dynamics.
\end{abstract}

% Center headings for table of contents, LOT, and LOF and make them smaller so
% that "Abstract", "Acknowledgments" and "Contents" all look alike. Comment out
% if you want the default. If you want more control, use the "tocloft" package.
\renewcommand{\contentsname}{\protect\centering\protect\Large Contents}
\renewcommand{\listtablename}{\protect\centering\protect\Large List of Tables}
\renewcommand{\listfigurename}{\protect\centering\protect\Large List of Figures}

\newcommand{\indep}{\mathrel{\text{\scalebox{1.07}{$\perp\mkern-10mu\perp$}}}}

\tableofcontents % Table of contents

% The rest of the front matter: Lists of tables, figures, dedication and
% acknowledment is optional. Comment out whatever you don't like
% \listoftables
\listoffigures
\begin{acknowledgments}
  TODO
\end{acknowledgments}
% \begin{dedication}
%   TODO
% \end{dedication}


%%%%%%%%%%%%%%%% MAIN BODY %%%%%%%%%%%%%%%%
\pagenumbering{arabic} % reset page numbering and switch to arabic

% Introductory chapter. Comment out if you don't have an intro chapter, but I
% think most committees expect you to have one.
% Don't number the intro chapter, but add to to the table of contents
\addcontentsline{toc}{chapter}{Introduction}
\chapter*{Introduction}\label{ch:intro}
\input{intro}


\chapter{Connected Worlds}\label{ch:1}
\input{chapter1}



\chapter{The Switching State Space Model}\label{ch:2}
\input{chapter2}



\chapter{User Study}\label{ch:3}
\input{chapter3}



\chapter{Conclusion}\label{ch:4}
\input{chapter4}

%%%%%%%%%%%%%%%% BACK MATTER %%%%%%%%%%%%%%%%

% Put appendices, bibliography, and supplemental materials here

% The bibliography may be single spaced within each entry, but must be
% double-spaced between each entry. Most bibliography styles leave space between
% entries, so that shouldn't be a problem.
\begin{singlespacing}
  % I like "References" better than "Bibliography"
  \renewcommand{\bibname}{References}

  % Any bibliohgraphy style that leaves space between entries is fine
  \bibliographystyle{ecca}
  \bibliography{references}
\end{singlespacing}

% Appendices from all chapters should go at the end
\input{appendix}

\end{document}
